---
title: "14 Inital survival Analysis"
author: "Katie M Susong"
date: "03-June-22"
---



Overview
========

Initial plots and correlation for survival/hatch rate.

Plots and Analysis
-------------------
  1. histogram of percent survival
         - the distribution is skewed to 0, given bounding at 0 and 1 pdf = beta
WITHIN THE TIRES 
  2. GDDs
      a. Plots 
            i. histogram
           ii. correlation scatter plots 
      b. Correlation tests : all measures show some correlation
            i.    GDDdjf10     : Highly correlated (0.906, p = 0.0001)
            ii.   GDDfull10    : Highly correlated ( 0.8827, p = 0.0003)
            iii.  GDDdjf -12   : Highly correlated (0.906, p = 0.0001)
            vi.   GDDfull -12  : Highly correlated ( 0.8827, p = 0.0003)
      c. Regression
            i.   null model: per.sur.live ~ 1 
            ii.  GDDdjf10 
            iii. GDDfull10
            vi.  GDDdjf -12
            v.   GDDfull -12
  3. Mean Jan T
      a. Plots 
      b. Correlation tests   : correlated (0.8269, p = 0.002)
      c. Regression
  4. Days below -12 
      a. Plots 
      b. Correlation tests   : correlated (-0.7635, p = 0.006) 
      c. Regression 
  5. Max Cons. Hrs below -12
      a. Plots 
      b. Correlation tests   : correlated (-0.7997, p = 0.0031)
      c. Regression 
  6. Min temperature
      a. Plots 
      b. Correlation tests   : correlated (0.7649, p = 0.0061)  
      c. Regression 
  7. Freeze-Thaw events
      a. Plots 
      b. Correlation tests   : not correlated (-0.2416, p = 0.4741)
      c. Regression 
      
OUTSIDE THE TIRE 
  8. Mean Jan ext. temperature
      a. plot
      b. Correlation test : correlated (0.7817952, p = 0.002665)
      c. Regression 
  9. external days below -12 
      a. plot
      b. Correlation test :  correlated (corr = -0.6952378, p = 0.01755)
      c. Regression - NA
  10. external 
      a. plot
      b. Correlation test : survival is  not correlated (corr = -0.6952378, p = 0.01755)
      c. Regression - NA
```{r Libraries, Functions, Data}
rm (list  = ls())
#libraries# 
library(dplyr)
library(ggplot2)
library(betareg)
library(sjPlot)
library(patchwork)

#### Functions ####
overdisp_fun <- function(model) {
  rdf <- df.residual(model)
  rp <- residuals(model,type="pearson")
  Pearson.chisq <- sum(rp^2)
  prat <- Pearson.chisq/rdf
  pval <- pchisq(Pearson.chisq, df=rdf, lower.tail=FALSE)
  c(chisq=Pearson.chisq,ratio=prat,rdf=rdf,p=pval)
}


#### Import data ####
setwd("~/Documents/CBS_PhD/Ae.albo_OW_2021/21.22_Ae.albopictus_Overwintering")
# import data avg. to the tire level 
tire <- read.csv("00_Data/21.22_bytire.hatch_temperature_summary_NO.Arl3.csv")
# format corrections
tire <- subset(tire, select = -c(X))
tire$site    <- factor(tire$site, levels = c("", "Car", "Bon", "Dek", "Arl", "Han", "Spo"))
tire$loc.id <- factor(tire$loc.id, levels = c("CTR",
                                              "IL_Car_T1", "IL_Car_T2",
                                              "IL_Bon_T1", "IL_Bon_T2",
                                              "IL_Dek_T1", "IL_Dek_T2",
                                              "WI_Arl_T1", "WI_Arl_T2",
                                              "WI_Han_T1", "WI_Han_T2",
                                              "WI_Spo_T1" ,"WI_Spo_T2"))

# import data avg. to the site
site <- read.csv("00_Data/21.22_bysite.hatch_temperature_summary_NO.Arl3.csv")
# format
site$site    <- factor(site$site, levels = c("CTR", "Car", "Bon", "Dek", "Arl", "Han", "Spo"))
site <- subset(site, select = -c(X))

# filter control out of df
site_n <- site %>%
  filter( site != "CTR") 
tire_n <- tire %>%
  filter( loc.id != "CTR")%>%
  filter( loc.id != "WI_Arl_T2")
```

### 1. histogram of percent survival
Positive skewed pdf. Large number of 0s. could be rare event zero inflation. (Ref. for 
zero inflation [here](https://espace.library.uq.edu.au/data/UQ_9053/tm_ele_8_05.pdf?Expires=1655991548&Key-Pair-Id=APKAJKNBJ4MJBJNC6NLQ&Signature=BFqhZraaKjEPZ-bfYlqtT7R2wdjr9rRiS3ZKp8Gc16QDLOOKaGW3o3btT5feWJALXUJnwfPRDWuwyv0qjHxHDB9UPCHGZMnUq0rUBBpBbQaDo2lukL~hEG2BA9v4nU1uDdrBbKb8S2-F8Re-4BeukFj9upWsoLwgJMprW0L32pJEdvZxE1q42AqhE1Ct3VzrtE9Qp1EJgzQQpaCUsTnF2DKOHmQBOXF2qX6N0qEkS1OITfivynX~S~1vrj0hYkYZF1qoEqmZV-OKnX7DOHSKj1-SbYJHB0G~e2W6V85dFGWKjDFzds2UiyexKcHTpQlNeh1FIuUIFGTOnUceK0Tbrw__) )
```{r 1. histogram of percent survival, echo=FALSE}
hist(tire$per.sur.live)
```

----------------------------------------------------------------
## Variable from WITH IN THE TIRES
### 2. GDDs v survival 
#### a. Plots
```{r GDDS histograms, echo=FALSE}
#' minimal distribution. high or low value
par(mfrow=c(2,2))
hist(tire$GDD_10_DJF)
hist(tire$GDD_n12_DJF)
hist(tire$GDD_10_FULL)
hist(tire$GDD_n12_DJF)
par(mfrow=c(1,1))
```
FIGURE NOTES:

 + Car (the site with the highest survival) has a much larger number of GDDs 
       regardless of metric
 + Over all there does not seem to be a strong trend for the corr between sur. and any GDDs
```{r 3a. Plots}
p1 <- ggplot(tire_n, aes(GDD_10_DJF, per.sur.live)) +
  geom_point(aes(col = site))+
  ggtitle("GDD_10_DJF")

p2 <- ggplot(tire, aes(GDD_n12_DJF, per.sur.live, col = site)) +
  geom_point() +
  ggtitle("GDD_n12_DJF")

p3 <- ggplot(tire, aes(GDD_10_FULL, per.sur.live, col = site)) +
  geom_point() +
  ggtitle("GDD_10_FULL")

p4 <- ggplot(tire, aes(GDD_n12_FULL, per.sur.live, col = site)) +
  geom_point() +
  ggtitle("GDD_n12_FULL")

p1 +p3 + p2+ p4
```


#### b.Correlation tests
CORRELATION TEST NOTE:
 
 + For all measurements of GDD: Per.sur is significantly correlated to the number of GDDS 

```{r GDDS Corr, echo=TRUE}
# Pearson correlation test for each GDDs metric
cor.test(tire_n$per.sur.live, tire_n$GDD_10_DJF, method = "pearson")
cor.test(tire_n$per.sur.live, tire_n$GDD_n12_DJF, method = "pearson")
cor.test(tire_n$per.sur.live, tire_n$GDD_10_FULL, method = "pearson")
cor.test(tire_n$per.sur.live, tire_n$GDD_n12_FULL, method = "pearson")
```

#### c. Regression
```{r}
# return to 2,2 plots of model test plots 
par(mfrow=c(2,2))
```

**Null Model**
```{r}
# null model 
null <- glm(per.sur.live ~ 1, data = tire)
summary(null)
```

**GDD (set to temperatures => 10 ºC) during DJF**
```{r}
GDDdjf10 <- glm(per.sur.live ~ GDD_10_DJF, data = tire_n) # GDD (djf 10ºC)
summary(GDDdjf10)
# measures of fitness 
plot(GDDdjf10)          # fitness plots
overdisp_fun(GDDdjf10)  # overdisp. function 
tab_model(GDDdjf10)
```
MODEL FITNESS

 + Residuals vs Fitted: even variance, homoscedasticity
 + Normal Q-Q:  1.0 + quarterly showing some non-linearity
 + Scale-Location: some pattern in the residuals, generally horizontal
 + Residual vs Leverage: Obs. 4 and 5 are influential
 + over dispersal: Not significantly overdisp. 

 **GDD (set to temperatures => 10 ??C) during full time**
```{r}
GDDfull10 <- glm(per.sur.live ~ GDD_10_FULL, data = tire_n) # GDD (full 10C)
summary(GDDfull10)
plot(GDDfull10)
overdisp_fun(GDDfull10)
```
MODEL FITNESS

 + Residuals vs Fitted: Some change in variance of residuals. Trend of residuals is noted
 + Normal Q-Q:  1.0 + Theo. Quantiles are not linear 
 + Scale-Location: strong pattern in residuals 
 + Residual vs Leverage: obs. 1 and 4 have strong leverage 
 + over dispersal: is not significantly overdisp.

**GDD (set to temperatures => -12 C) during DJF time**
```{r}
# GDD (full 10??C)
GDDdjfn12 <- glm(per.sur.live ~ GDD_n12_DJF, data = tire_n)
summary(GDDdjfn12)
plot(GDDdjfn12)
overdisp_fun(GDDdjfn12)
```
MODEL FITNESS

 + Residuals vs Fitted: limited heteroscedasticity
 + Normal Q-Q:  1.0 + Theo. Quantiles are not linear 
 + Scale-Location: mild trend in small predicted values
 + Residual vs Leverage:  obs. 4 and 5 have strong leverage
 + over dispersal: is not overdisp.

**GDD (set to temperatures => -12 C) during Full time**
```{r}
# GDD (full 10??C)
GDDfulln12 <- glm(per.sur.live ~ GDD_n12_FULL, data = tire_n)
summary(GDDfulln12)
plot(GDDfulln12)
overdisp_fun(GDDfulln12)

plot_model(GDDfulln12)
```
MODEL FITNESS

 + Residuals vs Fitted: trend in variance, possible heteroscad. 
 + Normal Q-Q:  1.0 + Theo. Quantiles are not linear 
 + Scale-Location: strong trend
 + Residual vs Leverage: obs. 1 and 4 have strong leverage
 + over dispersal: in not overdisp.

### 3. Survival v Mean Jan tire temperature 
#### a. Plots
```{r}
ggplot(tire, aes(JANmeanT, per.sur.live, col = site))+
  geom_point()
```
#### b. Correlation test
CORRELATION TEST NOTES:

 + Survival is strongly correlated to the internal tire temperature. Corr = 0.8269 (p=0.001)
 
```{r}
# Pearson correlation test
cor.test(tire_n$per.sur.live, tire_n$JANmeanT, method = "pearson")
```
#### c. Regression 
```{r}
reg.JanMeanT <- glm(per.sur.live ~ JANmeanT, data = tire_n)
# summary
summary(reg.JanMeanT)
plot(reg.JanMeanT)
overdisp_fun(reg.JanMeanT)
```
MODEL FITNESS

 + Residuals vs Fitted: generally even distribution
 + Normal Q-Q:  generally linear
 + Scale-Location: some pattern in the residuals, generally horizontal
 + Residual vs Leverage: Obs. 3 is influential
 + over dispersal: Not significantly overdisp. 
 
### 4. Survival v DaysB12T
#### a. Plots 
```{r}
ggplot(tire, aes(DaysB12T, per.sur.live, col = site))+
  geom_point()
```
#### b. Correlation test
CORRELATION TEST NOTES:

 + survival is strongly correlated to the number of Days below -12C ( Corr = -0.763, p = 0.006)
 
```{r}
# Pearson correlation test
cor.test(tire_n$per.sur.live, tire_n$DaysB12T, method = "pearson")
```
#### c. Regression 
```{r}
reg.DaysB12T <- glm(per.sur.live ~ DaysB12T, data = tire_n)
# summary
summary(reg.DaysB12T)
plot(reg.DaysB12T)
overdisp_fun(reg.DaysB12T)
```
MODEL FITNESS

 + Residuals vs Fitted: possible slight hetorscad. slight if present. 
 + Normal Q-Q:  generally linear
 + Scale-Location: some pattern in the residuals, upperward trend later in predicted values 
 + Residual vs Leverage: Obs. 3 is influential
 + over dispersal: Not significantly overdisp. 

### 5. Survival v hrB12Tcon
#### a, Plot 
```{r}
ggplot(tire_n[c(1:7, 9:12),], aes(MAXHrsB12conT, per.sur.live, col = site))+
  geom_point()
```
#### b. Correlation test
CORRELATION TEST NOTES:

 + survival is not significantly correlated to the number of MAX number of consecutive hours below  below -12C ( Corr = -0.5169, p = 0.0853)
 
```{r}
# Pearson correlation test
cor.test(tire_n[c(1:7, 9:12),]$per.sur.live, tire_n[c(1:7, 9:12),]$MAXHrsB12conT, method = "pearson")
```
#### c. Regression 
```{r}
reg.MAXHrsB12conT <- glm(per.sur.live ~ MAXHrsB12conT, data = tire_n[c(1:7, 9:12),])
# summary
summary(reg.MAXHrsB12conT )
plot(reg.MAXHrsB12conT )
overdisp_fun(reg.MAXHrsB12conT )
```
MODEL FITNESS

 + Residuals vs Fitted: heteroscad. 
 + Normal Q-Q:  generally linear, upward trend after 1.0 T.Qu
 + Scale-Location: pattern in the residuals, upward trend later in predicted values 
 + Residual vs Leverage: Obs. 3 and 8 is influential
 + over dispersal: Not significantly overdisp. 

### 6. Survival v MinT
#### a, Plot 
```{r}
ggplot(tire, aes(MinT, per.sur.live, col = site))+
  geom_point()
```
#### b. Correlation test
CORRELATION TEST NOTES:

 + survival is  significantly correlated to the minimum temperature (corr = 0.764, p = 0.0061)
 
```{r}
# Pearson correlation test
cor.test(tire_n$per.sur.live, tire_n$MinT, method = "pearson")
```
#### c. Regression 
```{r}
reg.MinT <- glm(per.sur.live ~ MinT, data = tire_n)
# summary
summary(reg.MAXHrsB12conT )
plot(reg.MAXHrsB12conT )
overdisp_fun(reg.MAXHrsB12conT )
```
MODEL FITNESS

 + Residuals vs Fitted: heteroscad. 
 + Normal Q-Q:  generally linear, upward trend after 1.0 T.Qu
 + Scale-Location: pattern in the residuals, upward trend later in predicted values 
 + Residual vs Leverage: Obs. 3 and 8 is influential
 + over dispersal: Not significantly overdisp. 

### 7. Survival v FreThaT
#### a, Plot 
```{r}
ggplot(tire_n[c(1:7, 9:12),], aes(FreThaT, per.sur.live, col = site))+
  geom_point()
```
#### b. Correlation test
CORRELATION TEST NOTES:

 + survival is not significantly correlated to the number of freeze thaw events( Corr = -0.242, p = 0.474)
 
```{r}
# Pearson correlation test
cor.test(tire_n[c(1:7, 9:12),]$per.sur.live, tire_n[c(1:7, 9:12),]$FreThaT, method = "pearson")
```
#### c. Regression 
```{r}
reg.FreThaT <- glm(per.sur.live ~ FreThaT, data = tire_n[c(1:7, 9:12),])
# summary
summary(reg.FreThaT )
plot(reg.MAXHrsB12conT )
overdisp_fun(reg.MAXHrsB12conT )
```
MODEL FITNESS

 + Residuals vs Fitted: heteroscad. 
 + Normal Q-Q:  generally linear, upward trend after 1.0 T.Qu
 + Scale-Location: pattern in the residuals, upward trend later in predicted values 
 + Residual vs Leverage: Obs. 3 and 8 is influential
 + over dispersal: Not significantly overdisp. 

## Variables from OUTSIDE THE TIRE
### 8. Survival v Mean Jan ext. temperature
#### a, Plot 
```{r}
ggplot(tire_n, aes(JANmeanA, per.sur.live, col = site))+
  geom_point()
```
#### b. Correlation test
CORRELATION TEST NOTES:

 + survival is  significantly correlated to the minimum temperature (corr = 0.764, p = 0.0061)
 
```{r}
# Pearson correlation test
cor.test(tire_n$per.sur.live, tire_n$JANmeanA, method = "pearson")
```
#### c. Regression 
```{r}
reg.JANmeanA <- glm(per.sur.live ~ JANmeanA, data = tire_n)
# summary
summary(reg.JANmeanA )
plot(reg.JANmeanA )
overdisp_fun(reg.JANmeanA )
```
MODEL FITNESS

 + Residuals vs Fitted: heteroscad. 
 + Normal Q-Q:  generally linear, upward trend after 1.0 T.Qu
 + Scale-Location: pattern in the residuals, upward trend later in predicted values 
 + Residual vs Leverage: Obs. 3 and 8 is influential
 + over dispersal: Not significantly overdisp. 

### 9. Survival v Days below -12??C
#### a. Plot
```{r}
ggplot(tire_n, aes(DaysB12A, per.sur.live, col = site))+
  geom_point()
```
#### b. Correlation test
CORRELATION TEST NOTES:

 + survival is  not correlated (corr = -0.6952378 
, p = 0.01755)
 
```{r}
# Pearson correlation test
cor.test(tire_n$per.sur.live, tire_n$DaysB12A, method = "pearson")
```
### 10. External Max Cons. Hrs below -12
#### a. Plot
```{r}
ggplot(tire_n, aes(MAXHrsB12conA, per.sur.live, col = site))+
  geom_point()
```
#### b. Correlation test
CORRELATION TEST NOTES:

 + survival is  significantly correlated to the minimum temperature (corr = 0.764, p = 0.0061)
 
```{r}
# Pearson correlation test
cor.test(tire_n$per.sur.live, tire_n$MAXHrsB12conA, method = "pearson")
```
